---
title: "FooDMe - Food DNA Metabarcoding - benchmarking report"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
theme: default
---

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}

knitr::opts_chunk$set(out.width = '80%',fig.asp= 0.5,fig.align='center',echo=FALSE, warning=FALSE, message=FALSE)
options(markdown.HTML.header = system.file("misc", "datatables.html", package = "knitr"))

library(DT, quietly = T)
library(tidyverse, quietly = T)

executor <- Sys.info()["user"]
```

---
author: `r paste0(executor)`
---

```{r init}
workdir <- snakemake@params[["workdir"]]

confmat <- snakemake@input[["confmat"]]
metrics <- snakemake@input[["metrics"]]
pr_curve <- snakemake@input[["pr_curve"]]
yields <- snakemake@input[["yields"]]
db <- snakemake@input[["db"]]
soft <- snakemake@input[["soft"]]

# infer run name from workdir
run <- basename(workdir)
#head(tail(strsplit(workdir,"/")[[1]],2),1)

# Number of samples
nsamples <- nrow(read.csv(file = yields, sep = "\t", check.names=FALSE))

# graphs
theme_set(theme_bw())
```

# Benchmarking report for run: `r run` {.tabset}

## Overview

* Run name: `r run`

* Number of samples: `r nsamples`

* Run directory: `r workdir`

* config file: <a href=../config.yaml>`r file.path(workdir,"config.yaml")`<a/>

* log files in folder: `r file.path(workdir,"logs")`

* snakemake logs in folder: `r file.path(workdir,".snakemake")`

* FooDMe version: `r snakemake@params[["version"]]`


## Yield

```{r yields}
data_table <- read.csv(file = yields, sep = "\t", check.names=FALSE)
datatable(data_table, filter = 'top', rownames = FALSE, escape = FALSE,
		extensions = list("ColReorder" = NULL, "Buttons" = NULL),
		options = list(	
					dom = 'BRrltpi',
					autoWidth=FALSE,
					scrollX = TRUE,
					lengthMenu = list(c(10, 50, -1), c('10', '50', 'All')),
					ColReorder = TRUE,
					buttons =
					list(
						'copy',
						'print',
						list(
						extend = 'collection',
						buttons = c('csv', 'excel', 'pdf'),
						text = 'Download'
						),
						I('colvis')
						)
						))
```



```{r yield_plots}
read_tsv(
        yields,
        col_names=c('sample', 'Total', 'Merged', 'Clustered', 'Assigned'),
        skip=1,
        show_col_types = FALSE) %>% 
    pivot_longer(!sample, names_to="Steps", values_to="Yield") %>% 
    mutate(Steps=fct_relevel(as_factor(Steps),'Total', 'Merged', 'Clustered', 'Assigned')) %>%
    ggplot(aes(x=Steps, y=Yield, fill=Steps))+
    geom_violin( scale = "count")+
    stat_summary(fun = median, geom="line", group=1)+
    stat_summary(fun = median, geom="point", group=1)+
    ylab("Yield [%]")
```

Step-wise yield distribution (violins) and median (black point) across samples.

## Metrics

```{r metrics}
data_table <- read.csv(file = metrics, sep = "\t", check.names=FALSE)
datatable(data_table, filter = 'top', rownames = FALSE, escape = FALSE,
		extensions = list("ColReorder" = NULL, "Buttons" = NULL),
		options = list(	
					dom = 'BRrltpi',
					autoWidth=FALSE,
					scrollX = TRUE,
					lengthMenu = list(c(10, 50, -1), c('10', '50', 'All')),
					ColReorder = TRUE,
					buttons =
					list(
						'copy',
						'print',
						list(
						extend = 'collection',
						buttons = c('csv', 'excel', 'pdf'),
						text = 'Download'
						),
						I('colvis')
						)
						))
```



```{r metrics_plot}
read_tsv("/home/debian/NGS/spezies_indev/FooDMe/.tests/benchmarking/metrics.tsv",
        show_col_types = FALSE) %>%
    filter(Sample!="aggregated") %>%
    pivot_longer(!Sample, names_to="Metric", values_to="value") %>%
    filter(Metric!='Distance') %>%
    filter(Metric!='Error') %>%
    mutate(value=value*100) %>%
    ggplot(aes(x=Metric, y=value, fill=Metric))+
    geom_violin(scale = "count")+
    stat_summary(fun = median, geom="line", group=1)+
    stat_summary(fun = median, geom="point", group=1)+
    ylab("Value [%]")
```

Classification metrics distribution (violins) and median (black point) across samples.
Global metrics from aggregated samples and qunatification are not displayed.



```{r error_plot}
read_tsv("/home/debian/NGS/spezies_indev/FooDMe/.tests/benchmarking/metrics.tsv",
        show_col_types = FALSE) %>%
    filter(Sample!="aggregated") %>%
    pivot_longer(!Sample, names_to="Metric", values_to="value") %>%
    filter(Metric=='Error') %>%
    mutate(value=value*100) %>%
    ggplot(aes(x=Metric, y=value, fill=Metric))+
    geom_violin(scale = "count")+
    stat_summary(fun = median, geom="point", group=1)+
    coord_flip()+
    ylab("Value [%]")
```

Mean realtive error distribtion (violins) and median (black point) across samples.



```{r distance_plot}
read_tsv("/home/debian/NGS/spezies_indev/FooDMe/.tests/benchmarking/metrics.tsv",
        show_col_types = FALSE) %>%
    filter(Sample!="aggregated") %>%
    pivot_longer(!Sample, names_to="Metric", values_to="value") %>%
    filter(Metric=='Distance') %>%
    ggplot(aes(x=Metric, y=value, fill=Metric))+
    geom_violin(scale = "count")+
    stat_summary(fun = median, geom="point", group=1)+
    coord_flip()+
    ylab("Value [%]")
```

Euclidian distance between predicted and expected values (violins) and median (black point) across samples.



```{r prcurve}
read_tsv("/home/debian/NGS/spezies_indev/FooDMe/.tests/benchmarking/pr_curve.tsv",
        show_col_types = FALSE) %>%
    arrange(Recall) %>%
    ggplot(aes(x=Recall, y=Precision))+
    geom_line()
```

Precision-Recall curve of aggregated results

## Confusion matrix

```{r confmat}
data_table <- read.csv(file = confmat, sep = "\t", check.names=FALSE)
datatable(data_table, filter = 'top', rownames = FALSE, escape = FALSE,
		extensions = list("ColReorder" = NULL, "Buttons" = NULL),
		options = list(	
					dom = 'BRrltpi',
					autoWidth=FALSE,
					scrollX = TRUE,
					lengthMenu = list(c(10, 50, -1), c('10', '50', 'All')),
					ColReorder = TRUE,
					buttons =
					list(
						'copy',
						'print',
						list(
						extend = 'collection',
						buttons = c('csv', 'excel', 'pdf'),
						text = 'Download'
						),
						I('colvis')
						)
						))
```

## Versions

### Databases

```{r db_versions}
db_table <- read.csv(file = db, sep = "\t", check.names=FALSE)
knitr::kable(db_table)
```

### Softwares

```{r soft_versions}
soft_table <- read.csv(file = soft, sep = "\t", check.names=FALSE)
knitr::kable(soft_table)
```

## Help

[Link to the repository](https://github.com/CVUA-RRW/FooDMe)


[Report a bug or ask a question](https://github.com/CVUA-RRW/FooDMe/issues/new)


[Last release](https://github.com/CVUA-RRW/FooDMe/releases)
